---
title: "conditionals_loops.qmd"
format: html
editor: visual
---
```{r}
library(dplyr)
library(stringr)
```

Task 1
Create an object called pm2_5 with a value of 48 (representing Particulate Matter 2.5, an indicator for air quality, in μgm3
 (see more about PM2.5 here).

Write an if - else if - else statement that returns “Low to moderate risk” if pm2_5 (for Particulate Matter 2.5) is less than 100, “Unhealthy for sensitive groups” if PM 2.5 is 100 <= pm2_5 < 150, and “Health risk present” if PM 2.5 is >= 150.

Test by changing the value of your pm2_5 object and re-running your statement to check.

```{r}
pm2_5 <- 48 
if (pm2_5 < 100){
  print("Low to moderate risk")
} else if (pm2_5 >= 100 & pm2_5<150 ){
  print("Unhealthy for sensitive groups")
} else if (pm2_5 >= 150){
  "Health risk present"
}


# whale_length_m<- 15
# if ( whale_length_m > 35) {
#   print("GIANT")
# }else if (whale_length_m >= 20 & whale_length_m<= 35){
#     print("MEDIUM")
#   }else if (whale_length_m <20) {print("SMALL")}

```
Task 2
Store the string “blue whale” as an object called species. Write an if statement that returns “You found a whale!” if the string “whale” is detected in species, otherwise return nothing. Test by changing the species string & re-running to see output.
```{r}
species <- "blue panda"
if (str_detect(species, "whale")) {
  print("You found a whale!")
} else { }
```

Task 3
Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredient with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:

A veggie burrito is the cost of a base burrito
A chicken burrito costs 3.00 more than a base burrito
A steak burrito costs 3.25 more than a base burrito

```{r}
base_burrito <- 6.5 
main_ingredient <- "chicken"
if (main_ingredient == "veggie"){
  print(base_burrito)
} else if (main_ingredient == "steak"){
  steakburrito <- base_burrito + 3.25
  print(steakburrito)
} else if (main_ingredient == "chicken"){
  chickenburrito <- base_burrito + 3.00
  print(chickenburrito)
}


```


Task 4
Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}
fish_counts <- c(8, 10, 12, 23)
fish_total <- sum(fish_counts)
fish_types <- c("goldfish", "tetras", "guppies", "mollies")
for (i in seq(fish_types)){
  print(paste("this type of fish:", fish_types[i], "is", fish_counts[i]/fish_total*100 , "percent of the tank"  ))
}

```
Task 5
There is an existing vector in R called month.name that contains all month names (just ry running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).
```{r}
for (i in seq(month.name)){
  print(paste(month.name[i], "is month", i))
}
```










